---
title: "AN588 Data Replication Assignment"
author: "Julianna D."
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: architect
    toc: yes
---
# Introduction
> I will be replicating a descriptive static, an inferential statistic, and a figure from the following paper: "Socioeconomic disparities in subway use and COVID-19 outcomes in New York City."  
>
> Article Citation: Karla Therese L Sy, Micaela E Martinez, Benjamin Rader, Laura F White, Socioeconomic Disparities in Subway Use and COVID-19 Outcomes in New York City, American Journal of Epidemiology, Volume 190, Issue 7, July 2021, Pages 1234–1242, https://doi.org/10.1093/aje/kwaa277
 
## Description of the Study
- briefly outline the goal of the original paper, the dataset used, and the analyses conducted
> The goal of the authors' research was to highlight how socioeconomic disparities in New York City were magnified during the COVID-19 pandemic during the onset of the stay-at-home mandaate. They looked at Subway data to assess mobility in different NYC boroughs  by looking at metro card swipes at different subway stations from January 2020 to April 2020 (Note: the lockdown began on March 15, 2020). While some residents chose to reduce their mobility on their own accord between January 2020 and April 2020, the authors found that 2 groups of residents remained mobile until the official lockdown in March. People of lower socioeconomic status who identified as non-White and people who were essential/health-care essential workers were more likely to continue using the Subway until the official lockdown mandate on March 15, 2020. 

## Dataset Used
> Dataset Citation: Sy, Karla Therese L.; Martinez, Micaela E.; Rader, Benjamin; White, Laura F. (2021). Socioeconomic disparities in subway use and COVID-19 outcomes in New York City [Dataset]. Dryad. https://doi.org/10.5061/dryad.vhhmgqnrh

## Analyses Conducted
> I will be replicating ___ for the descriptive statistic, Figure 4, and 95% confidence intervals for median income for the inferential statistic. 

# Getting Started

## Load Packages
```{r, warning=FALSE}
library(ggplot2) # for data visualization
library(dplyr) # for data manipulation
library(segmented) # for regression models
```

# 1. Descriptive Statistics

## Descriptive Table?

# 2. Inferential Statistics

## Import the Dataset
> Please download the following datasets from my repository folder "Sy open data": 3._Longitudinal_Mobility_zcta_borough.csv

```{r}
d <- read.csv("Sy open data/3._Longitudinal_Mobility_zcta_borough.csv", header=TRUE)
head(d)
x <- read.csv("Sy open data/1._Cross-sectional_COVID19_mobility_zcta.csv", header=TRUE)
head(x)
```

## 95% Confidence Interval for Median Income
```{r}
nyc_change_long2_zcta_nosi1 <- left_join(d, x[,c("ZCTA","median_income")], by=c("ZCTA"="ZCTA"))
nyc_change_long2_zcta_nosi11 <- nyc_change_long2_zcta_nosi1 %>% mutate(income_d = ntile(median_income, 5))

#To get the quantile minimum and maximum for each income group, to be used as labels in Figure 3

quantile <- nyc_change_long2_zcta_nosi11 %>%
  group_by(income_d) %>%
  summarise(min=min(median_income,na.rm=TRUE), max=max(median_income,na.rm=TRUE))
quantile
```

# 3. Replicating Figure 4
> This figure assesses the association between sociodemographic variables, subway use, and COVID-19 rate per 100,0000 population. 

## Import the Dataset
> Please download the following dataset from my repository folder "Sy open data": 5._Regression_output.csv

```{r}
regression_output <- read.csv("5._Regression_output.csv", stringsAsFactors = TRUE)
head(regression_output)
```


## Preparing the Data
> First, I'm going to rename the variables for the legend

```{r}
# Step 1: REnaming the variables from the dataset and making labels for the legend
regression_output$Variable <- factor(regression_output$Variable, levels = c("hs",
                                                "healthins",
                                                "minority",
                                                "essential",
                                                "health",
                                                "essential_both",
                                                "median_age",
                                                "Median_Income",
                                                "subway_use"),
                       labels = c("High school education",
                                  "Uninsured",
                                  "Non-White/Hispanic",
                                  "Non-health-care essential services",
                                  "Health-care essential services",
                                  "All essential services",
                                  ">75 Years of age",
                                  "Median income",
                                  "Mobility")) 
```

> Next, I want to separate the data by Outcome: "COVID-adjusted" and "Subway_Use"

```{r}
# Step 2: Separating the "Outcome" column into "COVID_adjusted" and "Subway_Use"
regression_output$Outcome <- factor(regression_output$Outcome, levels = c("COVID_adjusted","Subway_Use"),
                       labels = c("Rate of COVID-19 (per 100k)\n\nEstimate (adjusted for testing):\nRisk Ratio",
                                  "Mobility (10 Units)\n\nEstimate (unadjusted):\nbeta")) 

# Step 3: Making 2 separate data frames for each outcome.
mobil <- regression_output[1:8,] #rows 1-8 refers to Subway_Use, or mobility
RR <- regression_output[9:17,] #rows 9-17 refer to Covid_adjusted for the Risk Ratio

# Step 4: Establishing the intercepts 
mobil$intercept <- 0 #the null is 0
RR$intercept <- 1 #the null for risk ratio is 1
```

## Subway_Use Estimated Beta
> Then, we will make our figures for the estimated Beta for "Subway_Use" and the "COVID-adjusted" Risk Ratios.

```{r, warning=FALSE}
fig4_1 <- ggplot(data=mobil, aes(x = Variable,y = RiskRatio, ymin = LowerLimit, ymax = UpperLimit )) +
  scale_y_continuous(limits=c(-.5,1.5), breaks = c(-.5,0,.5,1,1.5)) +
  geom_pointrange(cex=0.3) +
  geom_hline(data = mobil, aes(colour=Variable, yintercept = intercept), linetype=2, show.legend = FALSE) + xlab('') + 
  ylab(expression(beta)) + geom_errorbar(aes(ymin=LowerLimit, ymax=UpperLimit), width=0.3,cex=0.7) +
  theme_classic(base_size=12) +
  coord_flip() + theme(legend.position="bottom") + guides(col=guide_legend(ncol=3, byrow=TRUE, reverse = TRUE, title.position = "bottom")) + theme(legend.text = element_text(size=12), text=element_text(size=12), axis.text=element_text(size=12, colour="black"), legend.margin = margin(10, 50, 10, 10)
)

plot(fig4_1, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE)
```

## Covid-Adjusted Risk Ratio
```{r}

fig4_2 <- ggplot(data=RR, aes(x = Variable,y = RiskRatio, ymin = LowerLimit, ymax = UpperLimit )) +
  scale_y_continuous(trans = "log", limits=c(.8,2.2), breaks = c(.8,1,1.2,1.4,1.6,1.8, 2.0, 2.2)) +
  geom_pointrange(cex=0.3) +
  geom_hline(data = RR, aes(colour=Variable, yintercept = intercept), linetype=2, show.legend = FALSE) + xlab('') + 
  ylab("Risk Ratio") + geom_errorbar(aes(ymin=LowerLimit, ymax=UpperLimit),width=0.3,cex=0.7) +
  theme_classic(base_size=12) +
  coord_flip() + theme(legend.position="bottom") + guides(col=guide_legend(ncol=3, byrow=TRUE, reverse = TRUE, title.position = "bottom")) + theme(legend.text = element_text(size=12), text=element_text(size=12), axis.text=element_text(size=12, colour="black"), legend.margin = margin(10, 50, 10, 10)
)

plot(fig4_2, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE)
```


# Sources
> https://stulp.gmw.rug.nl/ggplotworkshop/comparinggroupstatistics.html#mean-errorbar-plots  
>
> https://stackoverflow.com/questions/14069629/how-can-i-plot-data-with-confidence-intervals  
>
> https://ggplot2.tidyverse.org/reference/geom_abline.html  
>
> https://rdocumentation.org/packages/ggplot2/versions/0.9.0/topics/geom_hline  
>
> https://cran.r-project.org/web/packages/segmented/segmented.pdf  
