---
title: "AN588 Data Replication Assignment"
author: "Julianna D."
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: architect
    toc: yes
---
# Introduction
> The data analyses and figure I will be replicating is from the following paper: "Socioeconomic disparities in subway use and COVID-19 outcomes in New York City."  

> Citation: Karla Therese L Sy, Micaela E Martinez, Benjamin Rader, Laura F White, Socioeconomic Disparities in Subway Use and COVID-19 Outcomes in New York City, American Journal of Epidemiology, Volume 190, Issue 7, July 2021, Pages 1234–1242, https://doi.org/10.1093/aje/kwaa277

> I will be replicating a descriptive static, an inferential statistic, and a figure. 

# Getting Started

## Load Packages
```{r}
library(lubridate)
library(ggplot2) # for data visualization
library(dplyr) # for data manipulation
library(segmented) # for regression models
```

# 1. Descriptive Statistics

## Import the Dataset
> Please download the following dataset from my repository: 1._Cross-sectional_COVID19_mobility_zcta.csv

```{r}
data <- read.csv("_", stringsAsFactors = TRUE)
head(data)
```

## Calculate the __

```{r}

```


# 2. Inferential Statistics

# 3. Replicating Figure 4
> This figure assesses the association between sociodemographic variables, subway use, and COVID-19 rate per 100,0000 population. 

## Import the Dataset
> Please download the following dataset from my repository: 5._Regression_output.csv

```{r}
regression_output <- read.csv("5._Regression_output.csv", stringsAsFactors = TRUE)
head(regression_output)
```

## Preparing the Data
> First, I'm going to rename the variables for the legend

```{r}
# Step 1: Renaming the variables from the dataset for the legend
regression_output$Variable <- factor(regression_output$Variable, levels = c("hs",
                                                "healthins",
                                                "minority",
                                                "essential",
                                                "health",
                                                "essential_both",
                                                "median_age",
                                                "Median_Income",
                                                "subway_use"),
                       labels = c("High school education",
                                  "Uninsured",
                                  "Non-White/Hispanic",
                                  "Non-health-care essential services",
                                  "Health-care essential services",
                                  "All essential services",
                                  ">75 Years of age",
                                  "Median income",
                                  "Mobility")) 
```

> Next, I want to separate the data by Outcome: "COVID-adjusted" and "Subway_Use"

```{r}
# Step 2: Separating the "Outcome" column
regression_output$Outcome <- factor(regression_output$Outcome, levels = c("COVID_adjusted","Subway_Use"),
                       labels = c("Rate of COVID-19 (per 100k)\n\nEstimate (adjusted for testing):\nRisk Ratio",
                                  "Mobility (10 Units)\n\nEstimate (unadjusted):\nbeta")) 

# Step 3: Making 2 separate data frames for each outcome 
mobil <- regression_output[1:8,]
RR <- regression_output[9:17,]
# Step 4: Establishing the intercepts 
mobil$intercept <- 0 #the null is 0
RR$intercept <- 1 #the null for risk ratio is 1
```

## Subway_Use Estimated Beta
> Then, we will make our figures for the estimated Beta for "Subway_Use" and the "COVID-adjusted" Risk Ratios.

## Subway_Use Estimated Beta
```{r}
fig4_1 <- ggplot(data=mobil, aes(x = Variable,y = RiskRatio, ymin = LowerLimit, ymax = UpperLimit )) +
  scale_y_continuous(limits=c(-.5,1.5), breaks = c(-.5,0,.5,1,1.5)) +
  geom_pointrange(cex=0.3) +
  geom_hline(data = mobil, aes(colour=Variable, yintercept = intercept), linetype=2, show.legend = FALSE) + xlab('') + 
  ylab(expression(beta)) + geom_errorbar(aes(ymin=LowerLimit, ymax=UpperLimit), width=0.3,cex=0.7) +
  theme_classic(base_size=12) +
  coord_flip() + theme(legend.position="bottom") + guides(col=guide_legend(ncol=3, byrow=TRUE, reverse = TRUE, title.position = "bottom")) + theme(legend.text = element_text(size=12), text=element_text(size=12), axis.text=element_text(size=12, colour="black"), legend.margin = margin(10, 50, 10, 10)
)

plot(fig4_1, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE)
```

## Covid-Adjusted Risk Ratio
```{r}

fig4_2 <- ggplot(data=RR, aes(x = Variable,y = RiskRatio, ymin = LowerLimit, ymax = UpperLimit )) +
  scale_y_continuous(trans = "log", limits=c(.8,2.2), breaks = c(.8,1,1.2,1.4,1.6,1.8, 2.0, 2.2)) +
  geom_pointrange(cex=0.3) +
  geom_hline(data = RR, aes(colour=Variable, yintercept = intercept), linetype=2, show.legend = FALSE) + xlab('') + 
  ylab("Risk Ratio") + geom_errorbar(aes(ymin=LowerLimit, ymax=UpperLimit),width=0.3,cex=0.7) +
  theme_classic(base_size=12) +
  coord_flip() + theme(legend.position="bottom") + guides(col=guide_legend(ncol=3, byrow=TRUE, reverse = TRUE, title.position = "bottom")) + theme(legend.text = element_text(size=12), text=element_text(size=12), axis.text=element_text(size=12, colour="black"), legend.margin = margin(10, 50, 10, 10)
)

plot(fig4_2, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE)
```


# Sources
> https://stulp.gmw.rug.nl/ggplotworkshop/comparinggroupstatistics.html#mean-errorbar-plots  
> https://stackoverflow.com/questions/14069629/how-can-i-plot-data-with-confidence-intervals  
> https://ggplot2.tidyverse.org/reference/geom_abline.html  
> https://rdocumentation.org/packages/ggplot2/versions/0.9.0/topics/geom_hline  
> https://cran.r-project.org/web/packages/segmented/segmented.pdf  
